---
title: "BDA400 Assignment 2 -- Technical Analysis using R (Preliminary Stage)"
author: "Onuegbo Nnaemeka Philip"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
---

## 1. Project Configuration

Portfolio symbols tracked in `portfolio.txt`:

```{r portfolio}
symbols <- readLines("portfolio.txt")
symbols
```

## 2. Loading Stock Data

```{r load-data}
source("scripts/load_stock_data.R")
stocks <- load_stock_data(portfolio_file = "portfolio.txt", data_dir = "data", src = "csv")

cat("Symbols loaded:", paste(names(stocks), collapse = ", "), "\n")
cat("Rows per symbol:\n")
sapply(stocks, nrow)
```

Preview of the AAPL data frame (first 6 rows):

```{r preview}
head(stocks[["AAPL"]])
```

## 3. Computing Statistics

```{r stats}
source("scripts/calculate_statistics.R")

all_summaries <- data.frame()
all_ma <- list()

for (sym in names(stocks)) {
  result <- calculate_statistics(stocks[[sym]], symbol = sym, ma_periods = c(20, 50))
  all_summaries <- rbind(all_summaries, result$summary)
  all_ma[[sym]] <- result$moving_averages
}

knitr::kable(all_summaries, digits = 2,
             caption = "Descriptive statistics for each portfolio stock (Close price)")
```

### Cross-check (independent verification)

```{r crosscheck}
for (sym in names(stocks)) {
  close_vals <- stocks[[sym]]$Close
  n <- length(close_vals)
  manual_mean <- sum(close_vals) / n
  manual_sd <- sqrt(sum((close_vals - manual_mean)^2) / (n - 1))

  r_mean <- mean(close_vals)
  r_sd <- sd(close_vals)

  cat(sprintf("%-6s | R mean=%.4f  manual mean=%.4f  match=%s | R sd=%.4f  manual sd=%.4f  match=%s\n",
              sym, r_mean, manual_mean, isTRUE(all.equal(r_mean, manual_mean)),
              r_sd, manual_sd, isTRUE(all.equal(r_sd, manual_sd))))
}
```

Both the mean and standard deviation computed by R's built-in `mean()`/`sd()` functions
match an independent, manually-coded recomputation for every symbol, satisfying the
course's required verification pattern (primary calculation + independent cross-check).

## 4. Displaying Imported Data

```{r display, results='hold'}
source("scripts/display_output.R")

for (sym in names(stocks)) {
  display_latest_quote(stocks[[sym]], symbol = sym)
}
```

### Most recent 10 trading days -- AAPL

```{r display-table}
display_price_table(stocks[["AAPL"]], symbol = "AAPL", n = 10)
```

## 5. Visualizations

### 5.1 Closing price with 20-day / 50-day moving averages

```{r chart-line, fig.width=9, fig.height=5}
ma_df <- all_ma[["AAPL"]]
ma_cols <- grep("^SMA_", names(ma_df), value = TRUE)
plot(ma_df$Date, ma_df$Close, type = "l", col = "black", lwd = 1.2,
     xlab = "Date", ylab = "Price (USD)",
     main = "AAPL -- Closing Price with Moving Averages")
lines(ma_df$Date, ma_df$SMA_20, col = "blue", lwd = 1.5)
lines(ma_df$Date, ma_df$SMA_50, col = "red", lwd = 1.5)
legend("topleft", legend = c("Close", "SMA 20", "SMA 50"),
       col = c("black", "blue", "red"), lty = 1, lwd = 1.5)
```

### 5.2 OHLC candlestick chart (last 180 trading days)

```{r chart-candles, fig.width=9, fig.height=6}
library(quantmod)
xts_obj <- xts::xts(stocks[["AAPL"]][, c("Open","High","Low","Close","Volume")],
                     order.by = stocks[["AAPL"]]$Date)
colnames(xts_obj) <- paste0("AAPL", c(".Open",".High",".Low",".Close",".Volume"))
chartSeries(tail(xts_obj, 180), name = "AAPL", theme = chartTheme("white"),
            TA = "addSMA(n=20, col='blue'); addSMA(n=50, col='red'); addVo()")
```

## 6. Full Portfolio Statistics Table

```{r final-table}
display_statistics_table(all_summaries)
```

---

## Appendix A: Steps Completed Outside R (documented with screenshots in the Word report)

1. GitHub account created and `TechnicalAnalysis` repository set up.
2. R and RStudio installed locally.
3. `quantmod` and `TTR` packages installed via `install.packages()`.
4. `portfolio.txt` created with 5 stock symbols.

See `Nnaemeka_BDA400_A02_Report.docx` for the full step-by-step walkthrough with screenshots.
